---
title: "Power Analysis in R"
author: "David Ly"
date: August 11, 2018
output:
  html_document: null
  number_sections: yes
  pdf_document: default
  toc: yes
  toc_depth: 3
---

# Outline
Power analysis is an important aspect in experimental design. It allows us to determine the sample size required to detect an effect of a given size with a degree of confidence. It also allows us to determine the probability of detecting an affect of a given size with a given level of confidence, under sample size constraints.

"Typically, before you do an experiment you should perform a power analysis to estimate the number ofobservations you need to have a good chance of detecting the effect you are looking for." 

The idea is to minimize cost and time using these tests.

# Links
* [Bio Handbook](http://www.biostathandbook.com/power.html)
* [R in Action: Chapter 10](http://ceal.fing.uncu.edu.ar/industrial/TyHM/DOE/R_in_Action.pdf)

### Definitions
* **Parameters**: sample size, effect size, significance level, power
* **Effect** size: size of the difference between your Ho and Ha that you hope to detect.
* **Power**: probability of detecting a given effect size with a given sample size
* **Significance Level**: significance level wat which the test should be contacted at
* **Beta**, in a power analysis, is the probability of accepting the null hypothesis, even though it is false (a false negative), when the real difference is equal to the minimum effect size. 

In statistical terminology, the minimum deviation from your null hypothesis that you hope to detect to reject.


### Install "pwr" package to use power analysis formulas

```{r}
if(!require(pwr)) {install.packages("pwr")}
```

# Examples

### Power analysis for a binomial test

```{r}
p0 <- 0.75
p1 <- 0.78
h = ES.h(p0, p1) # Calculate effect size

pwr.p.test(
  h = h,
  n = NULL, # NULL tells the function to calculate this
  sig.level = 0.05,
  power = 0.90, #1 - beta (1-.1(Type 2 Error)) = 0.90 
  alternative = "two.sided"
)
```

```{r}
## Answer
# proportion power calculation for binomial distribution (arcsine transformation) 
# 
# h = 0.07078702
# n = 2096.953
# sig.level = 0.05
# power = 0.9

# alternative = two.sided
```

### Power Analysis for unpaired t-test

Add in cohen's d formula: An effect size used to indicate standardized between between two means 

Cohen's D: This can be used in t-test/ANOVA results (appropriate effect size for the comparison of two means)

```{r}
m1 = 66.6 # mean for sample 1
m2 = 64.6 # mean for sample 2
s1 = 4.8  # sd for sample 1
s2 = 3.6  # sd for sample 2

cohen.d <- (m1 - m2)/sqrt(((s1^2) + (s2^2))/2) 

pwr.t.test(
  n = NULL,
  d = cohen.d,
  sig.level = 0.05, # type 1 probability
  power = 0.80,     # 1 minus type 2 probability
  type = "two.sample",
  alternative = "two.sided"
)
```

```{r}
## Answer
# Two-sample t test power calculation

# n = 71.61288 (Approx. 72)
# d = 0.4714045
# sig.level = 0.05
# power = 0.8
# alternative = two.sided

# NOTE: n is number in *each* group
```

### Calculate sample size for ANOVA (comparison of more than 2 groups) to get power = 0.80

```{r}
pwr.anova.test(
  n = NULL,
  k = 5,
  f = 0.25, # effect size is f in ANOVA
  sig.level = 0.05,
  power = 0.80
)
```

```{r}
## Answer
# Balanced one-way analysis of variance power calculation 

# k = 5
# n = 39.1534 (Approx. 40)
# f = 0.25
# sig.level = 0.05
# power = 0.8

# NOTE: n is number in each group
```

### Question: Detect if there is a difference between group 1 vs. group 2. How many students are needed to sample to have an 80% chance of a difference this large to be significant.

```{r}
mm1  = 66.6  # Mean for sample 1
mm2  = 64.6  # Mean for sample 2
sd1  = 4.80  # Std dev for sample 1
sd2  = 3.60  # Std dev for sample 2

Cohen.d = (mm1 - mm2)/sqrt(((sd1^2) + (sd2^2))/2)  

pwr.t.test(
  n = NULL,
  d = Cohen.d,
  sig.level = 0.05,
  power = 0.80,
  type = "two.sample",
  alternative = "two.sided"
)
```

```{r}
## Answer
# The result is 72, meaning that if group 2 were really 2 inches shorter than group 1 students, you'd need at least 72 students in each class to detect a significant difference 80% of the time, if the true difference really is 2.0 inches.
```

```{r}
## Result Summary
# Two-sample t test power calculation 

# n = 71.61288
# d = 0.4714045
# sig.level = 0.05
# power = 0.8
# alternative = two.sided

# NOTE: n is number in *each* group
```


### Test for 90% power of test (Need at least 96 students from each group to detect a significant difference 90% of the time.)

```{r}
pwr.t.test(
  n = NULL,
  d = Cohen.d,
  sig.level = 0.05,
  power = 0.90,
  type = "two.sample",
  alternative = "two.sided"
)
```

```{r}
## Result
# Two-sample t test power calculation 

# n = 95.53742
# d = 0.4714045
# sig.level = 0.05
# power = 0.9
# alternative = two.sided

# NOTE: n is number in *each* group
```

